---
output:
  html_document:
    css: www/styles.css
    anchor_sections: FALSE
    includes:
      in_header: "header_manual.html"

params:
  year_curr: 2024 # Current year of sampling
  all_years: TRUE # If FALSE, only reports on year_curr; TRUE reports on all years; not enabled yet- currently all years
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = FALSE, vline = "|")
options(knitr.kable.NA = '', scipen = 100)
```

## MacroPlot and Sample Event checks (PCM) {.tabset .tabset-pills}
### MacroPlot Checks {.tabset}
```{r libs, include = F}
library(plantcomNGPN)
library(tidyverse) # dplyr, purrr, tidyr
library(knitr) # for kable and include_graphic()
library(kableExtra) # for custom kable features
library(sf) # for checking plot coords against park bounding boxes
library(data.table) # faster summarizing than dplyr for big datasets
library(DT) # much faster at making tables for large dfs than kable

all_years <- params$all_years
year_curr <- params$year_curr
year_range <- if(all_years == TRUE){2011:year_curr} else {year_curr}
year_hist <- 2011:(year_curr - 1)

```

```{r imports, include = F}
# If SQL setup:
# importData(type = 'local',
#            dbname = c("FFI_RA_AGFO", "FFI_RA_BADL", "FFI_RA_DETO", "FFI_RA_FOLA",
#                       "FFI_RA_FOUS", "FFI_RA_JECA", "FFI_RA_KNRI", #"FFI_RA_MNRR",
#                       "FFI_RA_MORU", "FFI_RA_SCBL", "FFI_RA_THRO", "FFI_RA_WICA"),
#            keep_tables = T)
#keep_tables = T, export_views = T, export_path = "C:/temp")

# If SQL not set up, use import below instead, and update the import_path:
importData(type = 'csv', 
           import_path = "C:/Users/kbailey/Documents/Development/plantcomNGPN/data/NGPN_FFI_table_export_20260408.zip",
           keep_tables = T) 
```

```{r source, include = F}
# source code to generate most of the datasets/checks that are then printed in the rmd.
source("NGPN_network_MacroPlot_QC_compile2.R")
#source("./docs/NGPN_network_QC_checks_compile.R")
```

#### Macroplot QA/QC Info {.tabset}
<span style="font-size:15px;">

<p><strong>Monitoring Status Base</strong></p>

<ul style="padding-left: 40px;">
  <li>This list shows the plots that need to have Monitoring status base added in each park. Monitoring status base needs to be filled in with <strong>PlantCommunity</strong>. To filter out the correct plots for analysis, this column needs to be populated.</li>
  
  <li>When adding the monitoring status base, please check to make sure all sampling protocols were collected. If any plots have a sampling protocol that were not collected, <strong>do not</strong> add monitoring status base.</li>
  
  <li>This list was created by taking all of the sample events from FFI, filtering with the panel-year schedule (panel_schedule.csv) to remove any off schedule samples, and selecting plots that did not have a monitoring status base.</li>
</ul>

<p><strong>Sampling Schedule Counts</strong></p>

<ul style="padding-left: 40px;">
  <li>This table shows all of the scheduled sampling events and the total number of plots sampled for each park.</li>
</ul>

</span>

#### Monitoring Status Base {.tabset}
```{r samp_events_base, eval = F}
df_base <- samp_events_base |> filter(park %in% parkcode) |> select(-park)
# yrs <- names(df_base[,grepl("yr", names(df_base))])

table_missing_base <- df_base |> 
  kable(format = 'html',
        align = 'c',
        caption = paste0('A table of all the NGPN PCM plots from **', 
                         parkcode, 
                         '** missing a monitoring status base. ',
                         'These need to be check and a monitoring status ',
                         'base added in FFI.')) |> 
  kable_styling(fixed_thead = T, 
                bootstrap_options = c("condensed", 
                                      "striped"), 
                full_width = T, 
                position = 'left', 
                font_size = 10) |> 
  column_spec(1:ncol(df_base),
              border_left = "1px solid grey", 
              border_right = "1px solid grey")

print(table_missing_base)
```

```{r, parktab, results = 'asis'}
for(i in seq_along(park_list_base)){
  parkcode = park_list_base[[i]]
  cat("##### ", parkcode, "{.tabset}", "\n\n")
  <<samp_events_base>>
  cat("\n\n")
  cat("\n\n")
}
```

#### Sampling Schedule Counts {.tabset}
```{r samp_events_count, eval = F}
dfms <- samp_events_count |> filter(park %in% parkcode) |> select(-park)
yrs <- names(dfms[,grepl("yr", names(dfms))])

plot_events_count <- dfms |> 
  kable(format = 'html',
        align = 'c',
        caption = paste0('Table 2. A table of NGPN PCM plots for **',
                         parkcode,
                         '**. Each value represents the number of plots sampled for each panel during the specified year.',
                         "<br>","<br>",
                         "The purpose is from the macroplot level. ",
                         "These plots were selected by filtering all data downloaded from FFI by the designated sampling schedule (panel_schedule.csv). ",
                         "Please check to make sure the nubmer of plots sampled maches with the number on the table.")) |> 
  kable_styling(fixed_thead = T, 
                bootstrap_options = c("condensed", 
                                      "striped"), 
                full_width = T, 
                position = 'left', 
                font_size = 10) |> 
  column_spec(1:ncol(dfms),
              border_left = "1px solid grey", 
              border_right = "1px solid grey")

print(plot_events_count)
```

```{r, parktab1, results = 'asis'}
for(i in seq_along(park_list_samp)){
  parkcode = park_list_samp[[i]]
  cat("##### ", parkcode, "{.tabset}", "\n\n")
  <<samp_events_count>>
  cat("\n\n")
  cat("\n\n")
}
```
