---
output:
  html_document:
    css: www/styles.css
    anchor_sections: FALSE
    includes:
      in_header: "header_manual.html"

params:
  year_curr: 2024 # Current year of sampling
  all_years: TRUE # If FALSE, only reports on year_curr; TRUE reports on all years; not enabled yet- currently all years
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = FALSE, vline = "|")
options(knitr.kable.NA = '', scipen = 100)
```

## MacroPlot and Sample Event checks (PCM) {.tabset .tabset-pills}
### MacroPlot Checks {.tabset}
```{r libs, include = F}
library(plantcomNGPN)
library(tidyverse) # dplyr, purrr, tidyr
library(knitr) # for kable and include_graphic()
library(kableExtra) # for custom kable features
library(sf) # for checking plot coords against park bounding boxes
library(data.table) # faster summarizing than dplyr for big datasets
library(DT) # much faster at making tables for large dfs than kable

all_years <- params$all_years
year_curr <- params$year_curr
year_range <- if(all_years == TRUE){2011:year_curr} else {year_curr}
year_hist <- 2011:(year_curr - 1)

```

```{r imports, include = F}
# If SQL setup:
# importData(type = 'local',
#            dbname = c("FFI_RA_AGFO", "FFI_RA_BADL", "FFI_RA_DETO", "FFI_RA_FOLA",
#                       "FFI_RA_FOUS", "FFI_RA_JECA", "FFI_RA_KNRI", #"FFI_RA_MNRR",
#                       "FFI_RA_MORU", "FFI_RA_SCBL", "FFI_RA_THRO", "FFI_RA_WICA"),
#            keep_tables = T)
#keep_tables = T, export_views = T, export_path = "C:/temp")

# If SQL not set up, use import below instead, and update the import_path:
importData(type = 'csv', 
           import_path = "C:/Users/kbailey/Documents/Development/plantcomNGPN/data/NGPN_FFI_table_export_20260408.zip",
           keep_tables = T) 
```

```{r source, include = F}
# source code to generate most of the datasets/checks that are then printed in the rmd.
source("NGPN_network_MacroPlot_QC_compile2.R")
#source("./docs/NGPN_network_QC_checks_compile.R")
```

#### Questions/Notes {.tabset}
<span style="font-size:15px;">
<ul><li> Monitoring Status Base - MS (PCM) </li>
  <ul>
<li> If there isn't anything in the MonitoringStatus_Base, does that row need to be dropped? Sometimes it looks like it should (e.g. AGFO_PCM_004) and sometimes it does (e.g. BADL_PCM_0005)</li>
<li> What does Dual mean in MonitoringStatus_Base? (e.g. DETO_PCM_005)</li>
<li> There are still quite a few samples that don't follow the 2-on, 3-off sampling protocol. Any ideas on what is going on there? (e.g. FOUS)</li>
<li> Is there anything that I might have dropped when filtering the dataset?</li>
  </ul>
</ul>

<ul><li> Monitoring Status Base - MS (FPCM) </li>
  <ul>
<li>There are some PCM plots scattered throughout. Are those because a fire came through and you sampled that plot? (e.g. AGFO_PCM_001) I selected those particular pcm plots because in MonitoringStatus_Base it contained a fire related designation.</li>
<li>Does anything look weirdly off? I know fire follows a pre-post burn sampling protocol, but I'm not sure if there is an easily identifiable patter similar to the pcm only plots.</li>
  </ul>
</ul>
</span>

#### Monitoring Status Base - MS (PCM) {.tabset}
```{r monstat_ms_park, eval = F}
dfms <- macro_samp_ms_pcm |> filter(park %in% parkcode) |> select(-park)
yrs <- names(dfms[,grepl("yr", names(dfms))])

# dfms$dup <- rowSums(dfms[,yrs] > 1, na.rm = T)


plot_mat_ms <- dfms |> 
  kable(format = 'html',
        align = 'c',
        caption = paste0('Table 1. Matrix of NGPN PCM and RCM plots for **',
                         parkcode,
                         '** by monitoring status base, purpose, and year sampled. This is for',
                         'NGPN plots only.',
                         "<br>","<br>",
                         "The default monitoring status base comes from the Monsitoring Status.",
                         "The purpose is from the macroplot level.",
                         "There isn't consisency in how either of these are labeled ",
                         "(e.g. AGFO has some without Monitoring Status Base, BADL sometimes has ",
                         "Panel1 in that column. What I need to know is whether the mismatches ",
                         "are okay, if there is some combination that I need to take out, or if ",
                         "there are plots missing.")) |> 
  kable_styling(fixed_thead = T, 
                bootstrap_options = c("condensed", 
                                      "striped"), 
                full_width = T, 
                position = 'left', 
                font_size = 10) |> 
  column_spec(1:ncol(dfms),
              border_left = "1px solid grey", 
              border_right = "1px solid grey")
  
# macro_samp_ms_pcm |> 
#   kable(format = 'html',
#         align = 'c',
#         caption = paste0('Table 1. matrix of NGP PCM and RCM plots for **')) |> 
#                          # ,
#                          # parkcode,
#                          # '** by monitoring status base, purpose, and year sampled')) |> 
#   kable_styling(fixed_thead = T, 
#                 bootstrap_options = c("condensed", 
#                                       "striped"), 
#                 full_width = T, 
#                 position = 'left', 
#                 font_size = 10) |> 
#   column_spec(1:ncol(macro_samp_ms_pcm),
#               border_left = "1px solid grey", 
#               border_right = "1px solid grey")
#   
#   column_spec(1:(ncol(dfms)-2), 
#               border_left = "1px solid grey", 
#               border_right = "1px solid grey")

# dfms <- macro_samp_ms |> filter(park %in% parkcode) |> select(-park)
# yrs <- names(dfms[,grepl("yr", names(dfms))])
# 
# dfms$dup <- rowSums(dfms[,yrs] > 1, na.rm = T)
# 
# plot_mat_ms <- dfms |> 
#   # select(-dup, -dup_ms) |> 
#   mutate(across(4:(ncol(dfms)-2), 
#          ~ cell_spec(.x, 
#                      background = case_when(is.na(.x) ~ "#ffffff",
#                                             .x == 1 ~ "#ffffff",
#                                             TRUE ~ "#F2F2A0")))) |> 
#   kbl(format = 'html', 
#       align = 'c',
#       caption = paste0('Table 1. Matrix of NGPN PCM plots for **', parkcode, '** by monitoring status_base, purpose, and year sampled. ',
#                        "Only plots with 'PCM' in their name are included. ",
#                        "Plots with LPCM in their name (legacy plots) are sorted last. ", "<br>","<br>",
#                        
#                        "The monitoring status in this table comes from the MonitoringStatus$MonitoringStatus_Base ", 
#                        "column, which has typos (eg '_' in the name) that make it difficult to use as a filter. ",
#                        "Typos are indicated by <b style='background-color:#ff9797;'>red</b>. ",
#                        
#                        "Plots with multiple rows have multiple monitoring statuses recorded over time, ",
#                        "and are color coded <b style='background-color:orange;'>orange</b>. ",
#                        "Are the different monitoring statuses correct/intentional? Does one indicate a NGPN sample event, ",
#                        "and another indicate a fire effects or other sample event? <br>", "<br>",
#                        
#                        
#                        "Values indicate the number of records for each plot, year, and monitoring status. ", 
#                        "Values > 1 indicate duplicate sampling events in the SampleEvent table for that year and ",
#                        "are highlighted <b style='background-color:#F2F2A0;'>yellow</b>. ",
#                        "Duplicates within a year mean there are multiple dates and SampleEvent_GUIDs for a given year,
#                        which is tricky to accommodate in table joins/analysis. It looks like these are plots that were 
#                        started on one day and finished on another. It would be better to just use the first day and
#                        SampleEvent in those cases.", "<br>","<br>",
#                        
#                        "Sampling schedule also does not exactly follow the 2 on 3 off, so Monitoring Status is not ",
#                        "the only way to pull out NGPN PCM monitoring events?"), escape = F) |> 
#   collapse_rows(1, valign = "top") |> 
#   kable_styling(fixed_thead = T, 
#                 bootstrap_options = c("condensed", 
#                                       "striped"), 
#                 full_width = T, 
#                 position = 'left', 
#                 font_size = 10) %>%
#   column_spec(1:(ncol(dfms)-2), 
#               border_left = "1px solid grey", 
#               border_right = "1px solid grey") 
# %>%
#   purrr::reduce(4:(ncol(dfms)-2), function(x, y){
#     col = dfms[,y]
#     cell_spec(x, y, "HTLM", background = case_when(is.na(dfms[,y]) ~ "#ffffff",
#                                              dfms[,y] == 1 ~ "#ffffff",
#                                              TRUE ~ "#F2F2A0"))}, .init = kbl(dfms)) %>%
#   #column_spec(1, background = ifelse(df$dup > 0, "#F2F2A0", "#ffffff")) %>%
#   column_spec(1, background = case_when(dfms$dup_ms > 0 ~ "orange",
#                                         dfms$dup > 0 ~ "#F2F2A0", 
#                                         TRUE ~ "white")) %>%
#   column_spec(3, background = ifelse(grepl("_|2011|2012|2014|2015|2016' '", dfms$MonitoringStatus_Base), "#ff9797", "white"))

print(plot_mat_ms)
```

```{r, parktab2, results = 'asis'}
for(i in seq_along(park_list)){
  parkcode = park_list[[i]]
  cat("##### ", parkcode, "{.tabset}", "\n\n")
  <<monstat_ms_park>>
  cat("\n\n")
  cat("\n\n")
}
```

#### Monitoring Status Base - MS (FPCM) {.tabset}
```{r monstat_ms_park_fpcm, eval = F}
dfms2 <- macro_samp_ms_fpcm |> filter(park %in% parkcode) |> select(-park)
yrs2 <- names(dfms2[,grepl("yr", names(dfms2))])

# dfms$dup <- rowSums(dfms[,yrs] > 1, na.rm = T)


plot_mat_ms_fpcm <- dfms2 |> 
  kable(format = 'html',
        align = 'c',
        caption = paste0('Table 1. matrix of NGP PCM and RCM plots for **',
                         parkcode,
                         '** by monitoring status base, purpose, and year sampled.',
                         'These plots are fire plots only.')) |> 
  kable_styling(fixed_thead = T, 
                bootstrap_options = c("condensed", 
                                      "striped"), 
                full_width = T, 
                position = 'left', 
                font_size = 10) |> 
  column_spec(1:ncol(dfms2),
              border_left = "1px solid grey", 
              border_right = "1px solid grey")

# macro_samp_ms_fpcm |>
#   kable(format = 'html',
#         align = 'c',
#         caption = paste0('Table 1. matrix of NGP PCM and RCM plots for **')) |>
#                          # ,
#                          # parkcode,
#                          # '** by monitoring status base, purpose, and year sampled')) |>
#   kable_styling(fixed_thead = T,
#                 bootstrap_options = c("condensed",
#                                       "striped"),
#                 full_width = T,
#                 position = 'left',
#                 font_size = 10) |>
#   column_spec(1:ncol(macro_samp_ms_fpcm),
#               border_left = "1px solid grey",
#               border_right = "1px solid grey")

print(plot_mat_ms_fpcm)
```

```{r, parktab3, results = 'asis'}
for(i in seq_along(park_list2)){
  parkcode = park_list2[[i]]
  cat("##### ", parkcode, "{.tabset}", "\n\n")
  <<monstat_ms_park_fpcm>>
  cat("\n\n")
  cat("\n\n")
}
```
