---
output:
  html_document:
    css: www/styles.css
    anchor_sections: FALSE
    includes:
      in_header: "header_manual.html"

params:
  year_curr: 2024 # Current year of sampling
  all_years: TRUE # If FALSE, only reports on year_curr; TRUE reports on all years; not enabled yet- currently all years
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = FALSE, vline = "|")
options(knitr.kable.NA = '', scipen = 100)
```

## MacroPlot and Sample Event checks {.tabset .tabset-pills}
### MacroPlot Checks {.tabset}

This document, checks the MarcoPlot and SampleEvents tables for data entry errors. This will check to make sure active plots have all the necessary information to correctly query. You will also be able to determine if the correct number of plots are being selected for analysis. 

```{r libs, include = F}
library(plantcomNGPN)
library(tidyverse) # dplyr, purrr, tidyr
library(knitr) # for kable and include_graphic()
library(kableExtra) # for custom kable features
library(sf) # for checking plot coords against park bounding boxes
library(data.table) # faster summarizing than dplyr for big datasets
library(DT) # much faster at making tables for large dfs than kable

all_years <- params$all_years
year_curr <- params$year_curr
year_range <- if(all_years == TRUE){2011:year_curr} else {year_curr}
year_hist <- 2011:(year_curr - 1)

```

```{r imports, include = F}
# If SQL setup:
# importData(type = 'local',
#            dbname = c("FFI_RA_AGFO", "FFI_RA_BADL", "FFI_RA_DETO", "FFI_RA_FOLA",
#                       "FFI_RA_FOUS", "FFI_RA_JECA", "FFI_RA_KNRI", #"FFI_RA_MNRR",
#                       "FFI_RA_MORU", "FFI_RA_SCBL", "FFI_RA_THRO", "FFI_RA_WICA"),
#            keep_tables = T)
#keep_tables = T, export_views = T, export_path = "C:/temp")

# If SQL not set up, use import below instead, and update the import_path:
importData(type = 'csv', 
           import_path = "C:/Users/kbailey/Documents/Development/plantcomNGPN/data/NGPN_FFI_table_export_20260408.zip",
           keep_tables = T) 

# PCM Panel sampling schedule
# Path will need to be updated
panel_sch_wide <- read.csv("C:/Users/kbailey/Documents/Development/plantcomNGPN/data/panel_schedule.csv",
                           na.strings = "")

# pivot to longer
panel_sch <- panel_sch_wide |>
  pivot_longer(!Year,
               names_to = "Panel") |>
  drop_na() |>
  # filtering to current date (will update every year)
  filter(Year <= as.integer(format(Sys.Date(), "%Y"))) |>
  select(Year,
         Panel)
```

```{r data_wrangling, include = F}
## Macroplot Checks ----
### Merging tables ----

#### Plot Matrix List
macro <- NGPN_tables$MacroPlot # list of plots and purpose
samp <- NGPN_tables$SampleEvent # list of sample events

#### formatting date
samp$SampleEvent_Date <-
  as.Date(substr(samp$SampleEvent_Date, 1, 11), format = "%Y-%m-%d")

#### selecting only pcm type plots
plots <- macro$MacroPlot_Name[grepl("_PCM_|_LPCM_|_FPCM_|_RCM_", macro$MacroPlot_Name)]
# plots <- macro$MacroPlot_Name[grepl("_PCM_|_RCM_", macro$MacroPlot_Name)]

### Macro plot df ----
macro_plots <- macro |>
  # filtering pcm plots
  filter(MacroPlot_Name %in% plots)|>
  # cleaning up columns
  select(MacroPlot_Name,
         MacroPlot_Purpose,
         MacroPlot_Type,
         MacroPlot_RegistrationUnit_GUID,
         MacroPlot_UTM_X,
         MacroPlot_UTM_Y,
         MacroPlot_UTMzone,
         MacroPlot_Datum,
         MacroPlot_DD_Lat,
         MacroPlot_DD_Long,
         MacroPlot_Elevation,
         MacroPlot_Aspect,
         MacroPlot_Azimuth,
         MacroPlot_SlopeHill,
         MacroPlot_SlopeTransect,
         MacroPlot_GUID) |>
  # naming parks
  mutate(park = substr(MacroPlot_Name, 1, 4)) |>
  # dropping duplicate plots
  distinct()

### Macro plot/sample events ----
# joining macro plots with sample events
macro_samp <- left_join(macro_plots, samp, by = c("MacroPlot_GUID" = "SampleEvent_Plot_GUID")) |>
  # cleaning up columns
  select(park,
         MacroPlot_Name,
         MacroPlot_GUID,
         MacroPlot_Purpose,
         SampleEvent_GUID,
         SampleEvent_Date,
         SampleEvent_DefaultMonitoringStatus) |>
  # removing duplicates
  distinct()

# adding year
macro_samp$year <- as.numeric(format(as.Date(macro_samp$SampleEvent_Date,
                                             format = "%Y-%m-%d"), "%Y"))

# NA to blanks
macro_samp$SampleEvent_DefaultMonitoringStatus[is.na(macro_samp$SampleEvent_DefaultMonitoringStatus)] <- "blank"

### Add monitoringstatus_base ----
monstat <- NGPN_tables$MonitoringStatus # monitoringstatus_base
mm_monstat_se <- NGPN_tables$MM_MonitoringStatus_SampleEvent # needed for GUID matching

# merge for correct GUID
macro_samp_ms1 <- left_join(macro_samp,
                            mm_monstat_se,
                            by = c("SampleEvent_GUID" = "MM_SampleEvent_GUID"))

# merge for monitoringstatus_base
macro_samp_ms2 <- left_join(macro_samp_ms1,
                            monstat,
                            by = c("MM_MonitoringStatus_GUID" = "MonitoringStatus_GUID",
                                   "datasource"))


macro_samp_ms <- macro_samp_ms2 |>
  # removing anything before 2011
  filter(year >= 2011) |>
  select(park,
         MacroPlot_Name,
         MonitoringStatus_Base,
         MacroPlot_Purpose,
         # SampleEvent_DefaultMonitoringStatus,
         SampleEvent_Date,
         year)

## Panel Filter ----
# removing sample events from macro_samp_ms that don't have a panel_yr match (keeping all observations for panel_sch)
samp_events_all <- right_join(macro_samp_ms,
                          panel_sch,
                          by = c("MacroPlot_Purpose" = "Panel",
                                 "year" = "Year"))

# filtering plots that NEED a monitoring status base
samp_events_base <- samp_events_all |>
  filter(is.na(MonitoringStatus_Base)) |>
  select(park,
         MacroPlot_Name,
         SampleEvent_Date,
         MonitoringStatus_Base) |>
  arrange(SampleEvent_Date, MacroPlot_Name)

# getting count for each panel-year
samp_events <- samp_events_all |>
  # filtering anything that isn't plant community samples
  mutate(keep = ifelse(grepl("PlantCommunity",
                             MonitoringStatus_Base), 1, 0)) |>
  filter(keep == 1) |> select(-keep) |>
  arrange(MacroPlot_Name, year)

# checking duplicate rows
sampe_events_dup <- samp_events[duplicated(samp_events) | duplicated(samp_events, fromLast = TRUE), ]

## Sample event counts ----
samp_events_count <- samp_events |>
  distinct() |>
  summarize(count = n(),
            .by = c(park,
                    MacroPlot_Purpose,
                    year)) |>
  pivot_wider(names_from = MacroPlot_Purpose,
              values_from = count) |>
  arrange(park, year)
```

```{r park_lists, include = F}
park_list_base <- sort(unique(samp_events_base$park))
park_list_samp <- sort(unique(samp_events_count$park))
```

```{r source, include = F}
# source code to generate most of the datasets/checks that are then printed in the rmd.
# source("NGPN_network_MacroPlot_QC_compile2.R")
#source("./docs/NGPN_network_QC_checks_compile.R")
```

#### Macroplot QA/QC Info {.tabset}
<span style="font-size:15px;">

<p><strong>Monitoring Status Base</strong></p>

<ul style="padding-left: 40px;">
  <li>This list shows the plots that need to have Monitoring status base added in each park. Monitoring status base needs to be filled in with <strong>PlantCommunity</strong>. To filter out the correct plots for analysis, this column needs to be populated.</li>
  
  <li>When adding the monitoring status base, please check to make sure all sampling protocols were collected. If any plots have a sampling protocol that were not collected, <strong>do not</strong> add monitoring status base.</li>
  
  <li>This list was created by taking all of the sample events from FFI, filtering with the panel-year schedule (panel_schedule.csv) to remove any off schedule samples, and selecting plots that did not have a monitoring status base.</li>
</ul>

<p><strong>Sampling Schedule Counts</strong></p>

<ul style="padding-left: 40px;">
  <li>This table shows all of the scheduled sampling events and the total number of plots sampled for each park.</li>
</ul>

</span>

#### Monitoring Status Base {.tabset}
```{r samp_events_base, eval = F}
df_base <- samp_events_base |> filter(park %in% parkcode) |> select(-park)
# yrs <- names(df_base[,grepl("yr", names(df_base))])

table_missing_base <- df_base |>
  kable(format = 'html',
        align = 'c',
        caption = paste0('A table of all the NGPN PCM plots from **',
                         parkcode,
                         '** missing a monitoring status base. ',
                         'These need to be check and a monitoring status ',
                         'base added in FFI.')) |>
  kable_styling(fixed_thead = T,
                bootstrap_options = c("condensed",
                                      "striped"),
                full_width = T,
                position = 'left',
                font_size = 10) |>
  column_spec(1:ncol(df_base),
              border_left = "1px solid grey",
              border_right = "1px solid grey")

print(table_missing_base)
```

```{r, parktab, results = 'asis'}
for(i in seq_along(park_list_base)){
  parkcode = park_list_base[[i]]
  cat("##### ", parkcode, "{.tabset}", "\n\n")
  <<samp_events_base>>
  cat("\n\n")
  cat("\n\n")
}
```

#### Sampling Schedule Counts {.tabset}
```{r samp_events_count, eval = F}
dfms <- samp_events_count |> filter(park %in% parkcode) |> select(-park)
yrs <- names(dfms[,grepl("yr", names(dfms))])

plot_events_count <- dfms |> 
  kable(format = 'html',
        align = 'c',
        caption = paste0('Table 2. A table of NGPN PCM plots for **',
                         parkcode,
                         '**. Each value represents the number of plots sampled for each panel during the specified year.',
                         "<br>","<br>",
                         "The purpose is from the macroplot level. ",
                         "These plots were selected by filtering all data downloaded from FFI by the designated sampling schedule (panel_schedule.csv). ",
                         "Please check to make sure the nubmer of plots sampled maches with the number on the table.")) |> 
  kable_styling(fixed_thead = T, 
                bootstrap_options = c("condensed", 
                                      "striped"), 
                full_width = T, 
                position = 'left', 
                font_size = 10) |> 
  column_spec(1:ncol(dfms),
              border_left = "1px solid grey", 
              border_right = "1px solid grey")

print(plot_events_count)
```

```{r, parktab1, results = 'asis'}
for(i in seq_along(park_list_samp)){
  parkcode = park_list_samp[[i]]
  cat("##### ", parkcode, "{.tabset}", "\n\n")
  <<samp_events_count>>
  cat("\n\n")
  cat("\n\n")
}
```
