---
output:
  html_document:
    css: www/styles.css
    anchor_sections: FALSE
    includes:
      in_header: "header_manual.html"

params:
  year_curr: 2024 # Current year of sampling
  all_years: TRUE # If FALSE, only reports on year_curr; TRUE reports on all years; not enabled yet- currently all years
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = FALSE, vline = "|")
options(knitr.kable.NA = '', scipen = 100)
```

### NGPN QC checks {.tabset}

```{r libs, include = F}
library(vegcomNGPN)
library(tidyverse) # dplyr, purrr, tidyr
library(knitr) # for kable and include_graphic()
library(kableExtra) # for custom kable features
```

```{r imports, include = F}
importData(type = 'local',
           dbname = c("FFI_RA_AGFO", "FFI_RA_BADL", "FFI_RA_DETO", "FFI_RA_FOLA",
                      "FFI_RA_FOUS", "FFI_RA_JECA", "FFI_RA_KNRI", "FFI_RA_MNRR",
                      "FFI_RA_MORU", "FFI_RA_SCBL", "FFI_RA_THRO", "FFI_RA_WICA"),
           export = F)
```

#### Plot List {.tabset}
WIP- add a table that shows the list of plots for each park, plus a year column showing when it was sampled (and if I think that was sampled as part of NGPN VS monitoring).
```{r}
macro <- NGPN_tables$MacroPlot
plots <- macro$MacroPlot_Name[grepl("_PCM_|_LPCM_|_FPCM_|_RCM_", macro$MacroPlot_Name)]

macro_plots <- macro |> mutate(park = substr(datasource, 8, 11)) |> 
  filter(MacroPlot_Name %in% plots)

macro2 <- left_join(macro_plots, NGPN_tables$MM_ProjectUnit_MacroPlot, 
                    by = c("MacroPlot_GUID" = "MM_MacroPlot_GUID", 'datasource'))
macroproj <- left_join(macro2, NGPN_tables$ProjectUnit, 
                       by = c("MM_ProjectUnit_GUID" = "ProjectUnit_GUID", "datasource"))

macroproj2 <- macroproj |> 
  mutate(park = substr(datasource, nchar(datasource)-3, nchar(datasource))) |> 
  select(park, MacroPlot_Name, MacroPlot_Purpose, MacroPlot_Type, ProjectUnit_Name, ProjectUnit_Agency, 
         MacroPlot_GUID) |> arrange(MacroPlot_Name)

macroproj_dups <- macroproj2 |> group_by(MacroPlot_Name, MacroPlot_GUID, MacroPlot_Purpose, ProjectUnit_Name) |> 
  summarize(num_rows = sum(!is.na(park)), .groups = 'drop') |> 
  arrange(ProjectUnit_Name) |> select(-MacroPlot_GUID) |> 
  pivot_wider(names_from = ProjectUnit_Name, values_from = num_rows) |> data.frame() |> 
  select(MacroPlot_Name, MacroPlot_Purpose, Park, everything()) |> 
  arrange(MacroPlot_Name)

macroproj_dups$num_recs <- apply(macroproj_dups[,4:ncol(macroproj_dups)], 1, function(x) sum(!is.na(x)))

macro_purp <- macro_plots |> select(park, MacroPlot_Name, MacroPlot_Purpose) |> 
  #unique() |> 
  mutate(pres = 1, 
         MacroPlot_Purpose = ifelse(is.na(MacroPlot_Purpose) | 
                                      MacroPlot_Purpose == "", "Unknown", MacroPlot_Purpose)) |>
  arrange(MacroPlot_Purpose) |> 
  pivot_wider(names_from = MacroPlot_Purpose, values_from = pres) |> 
  arrange(MacroPlot_Name)
```

#### Purpose {.tabset}
This table only includes plots with "_PCM_" in the plot name, which are assumed to be NGPN monitoring plots. 
```{r include = F}
kbl_purpose <- kable(macro_purp, format = 'html', align = 'c',
                     caption = 'NGPN PCM plots by purpose in MacroPlot table') |> 
  kable_styling(fixed_thead = T, bootstrap_options = c("condensed", "striped"), 
                full_width = T, position = 'left', font_size = 10) |> 
  column_spec(1:ncol(macro_purp), border_left = "1px solid grey", border_right = "1px solid grey")

```

```{r}
kbl_purpose
```

#### Project {.tabset}
```{r}
kbl_proj <- kable(macroproj_dups, format = "html", align = 'c',
                  caption = "NGPN PCM plots by project (strata) in MacroPlot table. Columns starting at Park indicate that a plot is in a given strata. Plots may be in multiple stratum.") |> 
  kable_styling(fixed_thead = T, bootstrap_options = c("condensed", "striped"),
                full_width = T, position = 'left', font_size = 10) |> 
  column_spec(1:ncol(macroproj_dups), border_left = "1px solid grey", border_right = "1px solid grey")

```

```{r}
kbl_proj
```

#### Macroplot Checks
```{r}
# Plots missing X-Y coordinates - usually they have lat/long. Would be helpful to fill in the UTMs too.
macro_vs <- getMacroPlot()

# Set bounding box for each park and check UTMs and/or lat/long against them:
# Plots with the wrong lat/long (WICA_FPCM_0165 and WICA_FPCM_0373 have lat/long swapped and UTM X/Y blank)
# BADL_FPCM_1724 has some screwy YCoordinates: 
# BADL_FPCM_1724

# THROS_PCM_0069 Macroplot_SlopeHill and _SlopeTransect has some 9999
```

#### Monitoring Status
```{r}
macro_guids <- unique(macro_plots$MacroPlot_GUID) # NGPN macroplot_guids for filter
mm_projunit <- NGPN_tables$MM_ProjectUnit_MacroPlot
regunit <- NGPN_tables$RegistrationUnit
projunit <- NGPN_tables$ProjectUnit
monstat <- NGPN_tables$MonitoringStatus
mm_monstat_se = NGPN_tables$MM_MonitoringStatus_SampleEvent
sampev <- NGPN_tables$SampleEvent |> filter(SampleEvent_Plot_GUID %in% macro_guids)

macro1 <- left_join(macro_plots, mm_projunit,
                    by = c("MacroPlot_GUID" = "MM_MacroPlot_GUID", "datasource"))
macro2 <- left_join(macro1, regunit, by = c("MacroPlot_RegistrationUnit_GUID" = "RegistrationUnit_GUID", "datasource"))
macro3 <- left_join(macro2, projunit,
                    by = c("MacroPlot_RegistrationUnit_GUID" = "ProjectUnit_RegistrationUnitGUID",
                           "MM_ProjectUnit_GUID" = "ProjectUnit_GUID",
                           "datasource")) |> unique()

mac_samp <- left_join(macro3, sampev, by = c("MacroPlot_GUID" = "SampleEvent_Plot_GUID", "datasource"), 
                      relationship = "many-to-many")
mac_samp_mm <- left_join(mac_samp, mm_monstat_se, by= c("SampleEvent_GUID" = "MM_SampleEvent_GUID", "datasource"),
                         relationship = 'many-to-many')
mac_samp_monstat <- left_join(mac_samp_mm, monstat,
                              by = c("MM_MonitoringStatus_GUID" = "MonitoringStatus_GUID",
                                     "datasource"))
mac_samp_monstat$SampleEvent_Date <-
  format(as.Date(mac_samp_monstat$SampleEvent_Date, format = "%Y-%m-%d %H:%m:%s"),
        "%Y-%m-%d")
mac_samp_monstat$year <- format(as.Date(mac_samp_monstat$SampleEvent_Date, format = "%Y-%m-%d"), "%Y")
mac_samp_monstat$month <- format(as.Date(mac_samp_monstat$SampleEvent_Date, format = "%Y-%m-%d"), "%m")
mac_samp_monstat$doy <- format(as.Date(mac_samp_monstat$SampleEvent_Date, format = "%Y-%m-%d"), "%j")

keep_cols <- c("MacroPlot_Name", "RegistrationUnit_Name", "MacroPlot_Purpose", "MacroPlot_Type", 
               #"ProjectUnit_Name", 
               "MacroPlot_UTM_X", "MacroPlot_UTM_Y",
               "MacroPlot_DD_Lat", "MacroPlot_DD_Long", "MacroPlot_Elevation",
               "MacroPlot_Aspect", "MacroPlot_Azimuth", "MacroPlot_SlopeHill", "MacroPlot_SlopeTransect",
               "SampleEvent_Date", "year", "month", "doy", #"SampleEvent_DefaultMonitoringStatus",
               "MonitoringStatus_Name", "MonitoringStatus_UV1",
               "MacroPlot_GUID", "SampleEvent_GUID", "MM_MonitoringStatus_GUID")

mac_samp_monstat2 <- mac_samp_monstat[,keep_cols]
mac_samp_monstat3 <- mac_samp_monstat2 |> filter(year >= 2011) |> 
  select(MacroPlot_Name, SampleEvent_Date, year, MonitoringStatus_Name) |> unique() |> 
  group_by(MacroPlot_Name, SampleEvent_Date, year, #ProjectUnit_Name, 
           MonitoringStatus_Name, #SampleEvent_DefaultMonitoringStatus,
           ) |> 
  summarize(num_samps = sum(!is.na(SampleEvent_Date)), .groups = 'drop')

mac_samp_monstat3$year_match <- 
  ifelse(substr(mac_samp_monstat3$MonitoringStatus_Name, 1, 4) %in% 2010:2024, 1, 0)

mac_samp_monstat4 <- mac_samp_monstat3 |> 
  filter(year_match == 1) |> 
  filter(nchar(MonitoringStatus_Name) > 4) |> 
  filter(!grepl("Other", MonitoringStatus_Name)) |> 
  group_by(MacroPlot_Name, SampleEvent_Date, year, MonitoringStatus_Name) |> 
  summarize(num_samps = sum(!is.na(year)), .groups = 'drop')  
  
mac_samp_monstat4$year_mismatch <- ifelse(
  mac_samp_monstat4$year != substr(mac_samp_monstat4$MonitoringStatus_Name, 1, 4), 1, 0)

monstat_yr_mismatch <- mac_samp_monstat4 |> filter(year_mismatch == 1) |> select(-num_samps, -year_mismatch, -monstat)

kbl_monstat_yr_mismatch <- kable(monstat_yr_mismatch, format = "html", align = 'c',
                  caption = "NGPN PCM plots with mismatch in year of SampleEvent_Date, and MonitoringStatus_Name.") |> 
  kable_styling(fixed_thead = T, bootstrap_options = c("condensed", "striped"),
                full_width = T, position = 'left', font_size = 10) |> 
  column_spec(1:ncol(monstat_yr_mismatch), border_left = "1px solid grey", border_right = "1px solid grey")

mac_samp_monstat4$monstat <- 
  substr(mac_samp_monstat4$MonitoringStatus_Name, 6, nchar(mac_samp_monstat4$MonitoringStatus_Name))

monstat_typo <- mac_samp_monstat4 |> group_by(MacroPlot_Name, monstat) |> 
  summarize(years = paste0(year, collapse = ", "),
            .groups = 'drop') |> 
  pivot_wider(names_from = monstat, values_from = years)

monstat_typo$inconsist <- rowSums(!is.na(monstat_typo[,2:ncol(monstat_typo)])) 
monstat_incon <- monstat_typo |> filter(inconsist > 1) |> arrange(MacroPlot_Name) |> 
  filter(inconsist > 1 | !is.na(`Plant Community`) | !is.na(Dual) | !is.na(PCM_Fire) |
           !is.na(Dual))

monstat_incon2 <- monstat_incon[,c("MacroPlot_Name", sort(names(monstat_incon[,2:(ncol(monstat_incon)-1)])))]

kbl_monstat_incon <- kable(monstat_incon2, format = "html", align = 'c',
                  caption = "NGPN PCM plots with inconsistently labeled MonitoringStatus_Name. Plots may be used for different monitoring purposes, but also seems some are incorrect. The years in the cell are years that a given monitoring status was recored (eg 2013 in PlantCommunity means there's a monitoring status name for that plot called '2013_PlantCommunity'.") |> 
  kable_styling(fixed_thead = T, bootstrap_options = c("condensed", "striped"),
                full_width = T, position = 'left', font_size = 10) |> 
  column_spec(1:ncol(monstat_incon2), border_left = "1px solid grey", border_right = "1px solid grey")

# incon_plot_list <- unique(monstat_incon2$MacroPlot_Name)
# 
# monstat_incon_full <- mac_samp_monstat4 |> filter(MacroPlot_Name %in% incon_plot_list) |> 
#   select(-num_samps) |> arrange(MacroPlot_Name, year, MonitoringStatus_Name)


```

#### Sample Event Checks
```{r}
kbl_monstat_yr_mismatch
```

```{r}
kbl_monstat_incon
```


```{r}
# Plots in MacroPlot table that don't have a corresponding SampleEvent:
#MORU_PCM_066; THRON_PCM_010; THRON_PCM_020; THRON_PCM_021; THRON_PCM_043; THROS_PCM_0029; THROS_PCM_0086

# MonitoringStatus_Name fixes: 14 plots in the MonitoringStatus table have a non-standard name: 
# 2009_Plant Community; 2018_Plant Community; 2024_Plant Community.

# Check the different module outputs in the package against the WICA data Dan sent.

# Check covpts_attr$Index has a bunch of blanks in it.

# the join to cover point attributes to cover point sample then sample events, causes many-to-many join in the sample events table. Need to figure out why.

# Plots with 2 visits within the same year: 
# AGFO_FPCM_067 in 2019; Same Date; MonitoringStatus_Name = 2019_PlantCommunity vs 2010_PlantCommunity. Same SampleEvent_GUID. I assume thw 2010 record is a mistake and will drop it from the getSampleEvent return by check that years match between dates and monitoring status. SampleEvent_GUID: 5F327BA4-006F-404A-AF37-05FEE4F202D5
# 
# JECA_PCM_134 in 2016; 2016-07-05; 2016-09-13; 2016_FirePlantCommunity; 2016_ForestStructure; 01yr02; 2016_ForestStructure; Based on other results from 2016 in JECA, 2016_ForestStructure appears to be the VS sample. 
# The FFI Data Depot returns 2 sets of CoverPoint data, and they're identical. 
# 
# JECA_PCM_038 in 2016; 2016-07-05; 2016-09-13; 2016_FirePlantCommunity; 2016_ForestStructure; 01yr02; 2016_ForestStructure;
# Also looks like sometimes MonitoringStatus_Name isn't consistent between ForestStructure or PlantCommunity across years, but usually within for the same plots. Then there's FirePlantCommunity, which appears to be different. 
```

