---
output:
  html_document:
    css: www/styles.css
    anchor_sections: FALSE
    includes:
      in_header: "header_manual.html"

params:
  year_curr: 2024 # Current year of sampling
  all_years: TRUE # If FALSE, only reports on year_curr; TRUE reports on all years; not enabled yet- currently all years
---

```{r setup, include = F}
knitr::opts_chunk$set(echo = FALSE, vline = "|")
options(knitr.kable.NA = '', scipen = 100)
```

## NGPN QC checks {.tabset .tabset-pills}

```{r libs, include = F}
library(plantcomNGPN)
library(tidyverse) # dplyr, purrr, tidyr
library(knitr) # for kable and include_graphic()
library(kableExtra) # for custom kable features
library(sf) # for checking plot coords against park bounding boxes
```

```{r imports, include = F}
importData(type = 'local',
           dbname = c("FFI_RA_AGFO", "FFI_RA_BADL", "FFI_RA_DETO", "FFI_RA_FOLA",
                      "FFI_RA_FOUS", "FFI_RA_JECA", "FFI_RA_KNRI", #"FFI_RA_MNRR",
                      "FFI_RA_MORU", "FFI_RA_SCBL", "FFI_RA_THRO", "FFI_RA_WICA"),
           export = F)
```

```{r include = F}
# source code to generate most of the datasets/checks that are then printed in the rmd.
source("NGPN_network_QC_checks_compile.R")
```

### Figuring out NGPN_PCM plots/visits {.tabset}
#### Questions/Notes {.tabset}
<span style="font-size:12px;">
<ul><li> In Monitoring Status Table 1, are the different monitoring statuses correct/intentional? </li>
  <ul>
<li> Is there a difference between AGFO_RCM_216's status of Riparian and RiparianCommunity, or should these be treated as the same thing? </li>
<li> In DETO, FOLA, JECA, KNRI, MORU, SCBL, THRO, and WICA: is there a difference between ForestStructure, PlantCommunity, and Dual?</li>
<li> AGFO_PCM_001 has 2 sample events in the SampleEvent table for year 2012. It appears that plot was started on 2012-06-04, and completed on 2012-06-07, and given 2 sample events as a result. If these are really part of the same sampling event, it would be better to assign all records to only one sample event for easier table joining. Can I assign the first sample event to all of the data for this plot in 2012? </li>
<li>KNRI_PCM_009 has 2 sample events in 2019 with the first date being 2019-07-24 and the 2nd being 2019-09-18. Both have no sample data associated with them and have notes about unsafe sampling conditions. Is there a column in the data that would help me filter these sample events out? </li>
<li>Sampling schedule of some plots does not exactly follow the 2 years on; 3 years off schedule. Is there a way to tell which visits are part of the GRTS sample design and which are not? Does that matter?</li>
  </ul>
</ul>
<ul>
<li>In Purpose Table 2, are the different monitoring statuses correct/intentional? </li>
  <ul>
  <li>Is purpose a way to distinguish between NGPN PCM plots and other plots? If so, are purposes beyond Panels and IM_Intensive considered part of NPGM PCM? Plots highlighted yellow are the ones in question.</li>
  </ul>
</ul>
<ul>
<li> In Project Table 3, Are plots highlighted yellow part of NGPN PCM?</li>
  <ul><li> Are there any plots that aren't Park = 1 but fall in other project/strata that should or should not be included if filtering on a strata other than park?
  </ul>
</ul>
</span>

#### Monitoring Status {.tabset}
```{r monstat_park, eval = F}
df <- macro_samp2 |> filter(park %in% parkcode) |> select(-park)
yrs <- names(df[,grepl("yr", names(df))])

df$dup <- rowSums(df[,yrs] > 1, na.rm = T)

plot_mat <- 
kable(df |> select(-dup, -dup_ms), format = 'html', align = 'c',
      caption = paste0('Table 1. Matrix of NGPN PCM plots for **', parkcode, '** by monitoring status and year sampled. ',
                       "Only plots with 'PCM' in their name are included. ",
                       "Plots with LPCM in their name (legacy plots) are sorted last. ", "<br>","<br>",
                       
                       "Plots with multiple rows have multiple monitoring statuses recorded over time, 
                       and are color coded <b style='background-color:orange;'>orange</b>. ",
                       "Are the different monitoring statuses correct/intentional? <br>", "<br>",
                       
                       "Values indicate the number of records for each plot, year, and monitoring status. ", 
                       "Values > 1 indicate duplicate sampling events in the SampleEvent table for that year and 
                       are highlighted <b style='background-color:#F2F2A0;'>yellow</b>. ",
                       "Duplicates within a year mean there are multiple dates and SampleEvent_GUIDs for a given year,
                       which is tricky to accommodate in table joins/analysis. It looks like these are plots that were 
                       started on one day and finished on another. It would be better to just use the first day and
                       SampleEvent in those cases.", "<br>","<br>",
                       
                       "Sampling schedule also does not exactly follow the 2 on 3 off, so Monitoring Status is not ",
                       "the only way to pull out NGPN PCM monitoring events?")) %>%
  kable_styling(fixed_thead = T, bootstrap_options = c("condensed", "striped"), 
                full_width = T, position = 'left', font_size = 10) %>%
  column_spec(1:(ncol(df)-2), border_left = "1px solid grey", border_right = "1px solid grey") %>%
  purrr::reduce(3:(ncol(df)-2), function(x, y){
    col = df[,y]
    column_spec(x, y, background = case_when(is.na(df[,y]) ~ "#ffffff",
                                             df[,y] == 1 ~ "#ffffff",
                                             TRUE ~ "#F2F2A0"))}, .init = .) %>%
  #column_spec(1, background = ifelse(df$dup > 0, "#F2F2A0", "#ffffff")) %>%
  column_spec(1, background = case_when(df$dup_ms > 0 ~ "orange",
                                        df$dup > 0 ~ "#F2F2A0", 
                                        TRUE ~ "white")) %>%
  collapse_rows(1, valign = "top")

print(plot_mat)
```

```{r, results = 'asis'}
for(i in seq_along(park_list)){
  parkcode = park_list[[i]]
  cat("##### ", parkcode, "{.tabset}", "\n\n")
  <<monstat_park>>
  cat("\n\n")
  cat("\n\n")
}
```

#### Purpose {.tabset}
```{r purpose_park, include = F, eval = F}
df2 <- macro_purp |> mutate(park = substr(MacroPlot_Name, 1, 4)) |> filter(park == parkcode) |> select(-park)
nonvs <- names(macro_purp[!grepl("Panel|IM_Intensive", names(macro_purp))])
df2$nonvs <- rowSums(!is.na(df2[,nonvs]))
df3 <- df2 |> arrange(nonvs, MacroPlot_Name)

kbl_purpose <- kable(df3 |> select(-nonvs), 
                     format = 'html', align = 'c',
                     caption = paste0('Table 2. NGPN PCM plots by purpose in MacroPlot table. ', 
                                      'Only includes plots with "PCM" in the plot name, ',
                                      'which are assumed to be NGPN plant community monitoring plots.',
                                      " Plots that with a purpose other than Panel# or IM_Intensive are 
                                      highlighted <b style='background-color:#F2F2A0;'>yellow</b> 
                                      and sorted at the bottom.", 
                                      " Are the yellow plots NPGN plant community monitoring plots?", "<br>","<br>",
                                      " Unknown purpose indicates that field is blank for a given plot.",
                                      " Can the combination of Monitoring Status and Purpose be used to pull out",
                                      " NGPN PCM sampling events?")) %>% 
  kable_styling(fixed_thead = T, bootstrap_options = c("condensed", "striped"), 
                full_width = F, position = 'left', font_size = 10) %>%
  purrr::reduce(14:(ncol(df3)-1), function(x, y){
    col = df3[,y]
    column_spec(x, y, background = ifelse(is.na(col), "#ffffff", "#F2F2A0"))}, .init = .) %>%
  column_spec(1, background = ifelse(df3$nonvs == 1, "#ffffff", "#F2F2A0")) %>%
  column_spec(1:(ncol(df3)-1), border_left = "1px solid grey", border_right = "1px solid grey") #%>%
  #scroll_box(width = "500px", height = "800px", fixed_thead = TRUE)

print(kbl_purpose)
```


```{r, results = 'asis'}
for(i in seq_along(park_list)){
  parkcode = park_list[[i]]
  cat("##### ", parkcode, "{.tabset}", "\n\n")
  <<purpose_park>>
  cat("\n\n")
  cat("\n\n")
}
```

#### Project {.tabset}
```{r proj_tbl, eval = F}
df4 <- macroproj_dups |> filter(park %in% parkcode) |> select(-park) |> arrange(desc(nonvs), MacroPlot_Name)

kbl_proj <- kable(df4 |> select(-num_recs, -nonvs), format = "html", align = 'c',
                  caption = "Table 3. NGPN PCM plots by project (strata) and purpose in the MacroPlot table. 
                  Columns starting at Park indicate that a plot is in a given strata. 
                  Plots may be in multiple stratum. Plots highlighted in <b style='background-color:#F2F2A0;'>yellow</b>
                  have a MacroPlot_Purpose that I'm unclear if these are NGPM PCM plots and are sorted at the bottom. 
                  Also, should plots with Project = 'IN_ACTIVE' be dropped?") %>%
  kable_styling(fixed_thead = T, bootstrap_options = c("condensed", "striped"),
                full_width = T, position = 'left', font_size = 10) %>%
  column_spec(1:2, background = ifelse(df4$nonvs == 1, "#ffffff", "#F2F2A0")) %>%
  column_spec(1:(ncol(df4)-2), border_left = "1px solid grey", border_right = "1px solid grey")

print(kbl_proj)

```

```{r, results = 'asis'}
for(i in seq_along(park_list)){
  parkcode = park_list[[i]]
  cat("##### ", parkcode, "{.tabset}", "\n\n")
  <<proj_tbl>>
    cat("\n\n")
  cat("\n\n")
}
```

### QC Check Table {.tabset}
```{r}
QC_check_table
```

```{r eval = macro_include, results = 'asis'}
cat("### MacroPlot Checks {.tabset}", "\n\n")
```

```{r eval = macro_pi_include, results = 'asis'}
cat("#### Plot Info {.tabset}", "\n\n")

check_null_print(kbl_macro_miss_utm, 5, "Missing UTM")

check_null_print(kbl_out_pts_utm, 5, "Incorrect UTM")

check_null_print(kbl_out_pts_dd, 5, "Incorrect DD")

cat("\n\n")
```


```{r eval = sampev_include, results = 'asis'}
cat("### Sample Event Checks {.tabset}", "\n\n")
```

```{r eval = sampev_ms_include, results = 'asis'}
cat("#### General {.tabset}", "\n\n")

check_null_print(kbl_miss_samp, 5, "Year Mismatch")

cat("\n\n")
```


```{r eval = sampev_ms_include, results = 'asis'}
cat("#### Monitoring Status {.tabset}", "\n\n")

check_null_print(kbl_monstat_yr_mismatch, 5, "Year Mismatch")

check_null_print(kbl_monstat_incon, 5, "Name Inconsistencies")

cat("\n\n")
```


